---
title: "Final Project"
output: html_notebook
---

Read in data 
```{r}
data_orgin = read.table('Ghosal2020.txt',sep = "\t",header = 1)
data_orgin
india = data_orgin[16,2:6]
india2 = data_orgin[16,3:5]
india3 = data_orgin[16,c(3,6)]
data = data_orgin[1:15,]
result = data[,7]
result
```

```{r}
fit <- lm(Week5deaths ~ Totalcases + Activecases +Recoverycases + Week4deaths + CFR, data = data)
summary(fit)
vif(fit)
```
From the coefficient, we can see that the week 5 death is largely determined by CFC. Also, we found out that the week 5 death is negatively accosiated with recovery cases, which is reasonable. 
```{r}
pairs(data[,2:7])
```
From the pairplot, we also can see that week5 death is slightly positive associated with total case, active case, whereas the week5 death is positively associated with CFC. (CFC is calculated based on other info)
```{r}
plot(fit$residuals, main = 'Residual plot', ylab = 'residuals')
abline(0,0,col = 'red')

qqnorm(fit$residuals)
qqline(fit$residuals, col="red")
```
```{r}
library(car)
avPlots(fit)
```


### Prediction
We want to predict week 5 death for india.
prediction is 195.6021
confidence interval for mean of predictions is [34.3962,356.808]
```{r}
predict(fit, newdata = data[,2:6])
predict(fit, newdata = india)
predict(fit, newdata = india,interval = "confidence")
predict(fit, newdata = india,interval = "prediction")
```

### Outlier
We did found outliers inside our dataset, which is China and Italy. We will analyze our fitting model by dropping info from china and italy, and then seperate 
```{r}
dropped= data[3:15,]
drop_china = data[-1,]
drop_italy= data[-2,]
```

```{r}
fitnew <- lm(Week5deaths ~ Totalcases + Activecases +Recoverycases + Week4deaths + CFR, data = dropped)
summary(fitnew)
pairs(dropped[,2:7])
plot(fitnew$residuals, main = 'Residual plot', ylab = 'residuals')
abline(0,0,col = 'red')
qqnorm(fitnew$residuals)
qqline(fitnew$residuals, col="red")
```
```{r}
fitchina <- lm(Week5deaths ~ Totalcases + Activecases +Recoverycases + Week4deaths + CFR, data = drop_china)
summary(fitchina)
pairs(drop_china[,2:7])
plot(fitchina$residuals, main = 'Residual plot', ylab = 'residuals')
abline(0,0,col = 'red')
qqnorm(fitchina$residuals)
qqline(fitchina$residuals, col="red")
```
```{r}
fitita <- lm(Week5deaths ~ Totalcases + Activecases +Recoverycases + Week4deaths + CFR, data = drop_italy)
summary(fitnew)
pairs(drop_italy[,2:7])
plot(fitita$residuals, main = 'Residual plot', ylab = 'residuals')
abline(0,0,col = 'red')
qqnorm(fitita$residuals)
qqline(fitita$residuals, col="red")
```

Although r-square did not incrase much, the pair plot did correlate better. However, we found out the slope all some variables changed a lot, especially CFR, it turned from positive 30 to -5 for both counries dropped, 20 for china dropped and 5 for only italy dropped. We doubt our original linear regression.  

## Scenario2
As we can see in the first part, the patterns between each pair of variables are not linear. Also, the residual plot shows extreme outliers, which means the linear model is very sensitive to outliers. These extreme outliers are influential and may cause errors to our linear model. Therefore, we revise the model by taking a logarithmic transformation of the entire dataset to reduce the effect of some influential skewed data(outliers).

But before we apply the logarithmic transformation of the dataset, we need first eliminate Brazil's data. That is because 'Week4deaths' and 'CFR' are both 0. If we take log of them, the entries will become '-inf' and cause errors.
```{r}
logdata<-data[-13,]
logdata<-data.frame(logdata['Countries'],log(logdata['Totalcases']),log(logdata['Activecases']),log(logdata['Recoverycases']),log(logdata['Week4deaths']),log(logdata['CFR']),log(logdata['Week5deaths']))
logdata
```
We draw the pair graph to see the pattern between each variable. Compare to the original pair graph, our new graph shows a noticeable linear relationship between each pair of variables.
```{r}
pairs(logdata[c("Totalcases", "Activecases", "Recoverycases","Week4deaths","CFR","Week5deaths")])
```
# fit
We use our new dataset to fit a new linear model. Although R-square decreases a little bit, but it still shows good fit. 
```{r}
fit2 <- lm(Week5deaths~Totalcases+Activecases+Recoverycases+Week4deaths+CFR, data = logdata)
summary(fit2)
vif(fit2)
```

# regression diagnostics
Now we do regression diagnostics of our new model.
By comparing the qqplots for the first model and the second model, we can see our new model's residuals lie approximately along the line which indicates a better fit.
Our new leverage plot also shows that our new model is less affected by outliers.
```{r}
qqPlot(fit, main = "QQ Plot")
qqPlot(fit2, main="QQ Plot") #qq plot for studentized resid
leveragePlots(fit) # leverage plots
leveragePlots(fit2)
```
Moreover, we have less residuals that are outside of Cook's distance, which indicate that we have less influential outliers.
However, in analysis of normality of residuals. We find that the first model's residuals are more likely to follow a normal distribution.
```{r}
plot(fit)
plot(fit2)
```

# prediction
Now, we use our new model to give a point estimate and a prediction interval for the number of deaths in India in week 5.
We estimate the week5 death to be 93.87909 with 
95% confidence interval to be [21.7043,406.0616]
95% prediction interval to be [12.76397,690.4813]
```{r}
exp(predict(fit2,log(india),interval = "confidence"))
exp(predict(fit2,log(india),interval = "prediction"))
```

The estimation using the second model differs from the first one.
Scenerio 3
```{r}

```

Scenerio 4
Then we use model selection to check whether 

# Model Selection

```{r}
set.seed(2020)
k <- 5
n <- 14
val.size <- floor(n/k)
folds_i <- sample(rep(1:k, length = n))
cv.mse <- rep(0, k)
ind.remain=1:n
for (round in 1:k){
  #val.ind <- sample(ind.remain, val.size, replace = FALSE)
  val.ind <- which(folds_i == round)
  fit<- lm(Week5deaths~Totalcases+Activecases+Recoverycases+Week4deaths+CFR, data = logdata[-val.ind,])
  y.hat <- predict(fit, logdata[val.ind,-7])
  cv.mse[round] <- mean((logdata[val.ind,7] - y.hat)^2)
}
mean(cv.mse)
```


```{r}
library(MASS)
library(datasets)
library(olsrr)
ols_step_backward_aic(fit, details = TRUE)
```

```{r}
ols_step_backward_aic(fit2, details = TRUE)
```


```{r}
library(MASS)
step1 <- step(fit, direction="backward")
step1$anova
```


```{r}
library(MASS)
step2 <- step(fit2, direction="backward")
step2$anova
```

```{r}
fit3 <- lm(Week5deaths~Totalcases + Activecases +Recoverycases + Week4deaths + CFR, data = data)#0.9701
#first iteration delete totalcases
fit3 <- lm(Week5deaths~Activecases +Recoverycases + Week4deaths + CFR, data = data)#0.9727
fit3 <- lm(Week5deaths~Totalcases + Recoverycases + Week4deaths + CFR, data = data)#0.9712
fit3 <- lm(Week5deaths~Totalcases + Activecases+Week4deaths + CFR, data = data)#0.9708
fit3 <- lm(Week5deaths~Totalcases + Activecases +Recoverycases + CFR, data = data)#0.8991
fit3 <- lm(Week5deaths~Totalcases + Activecases +Recoverycases + Week4deaths, data = data)#0.9715
summary(fit3)
#second iteration (delete CFR
fit3 <- lm(Week5deaths~Recoverycases + Week4deaths + CFR, data = data)#0.9701
fit3 <- lm(Week5deaths~Activecases + Week4deaths + CFR, data = data)#0.945 
fit3 <- lm(Week5deaths~Activecases +Recoverycases + CFR, data = data)#0.8389
fit3 <- lm(Week5deaths~Activecases +Recoverycases + Week4deaths, data = data)#0.974 use this
#third iteration delete
fit3 <- lm(Week5deaths~Recoverycases + Week4deaths, data = data)#0.9723
fit3 <- lm(Week5deaths~Activecases + Week4deaths, data = data)#0.9491
fit3 <- lm(Week5deaths~Activecases +Recoverycases, data = data)#0.7353
summary(fit3)
#decide to delete totalcases and CFR
fit.ms1 <- lm(Week5deaths~Activecases +Recoverycases + Week4deaths, data = data)
vif(fit.ms1)
summary(fit.ms1)
#predict
predict(fit.ms1,india,interval = "confidence")
predict(fit.ms1,india,interval = "prediction")

```


```{r}
fit4 <- lm(Week5deaths~Totalcases + Activecases +Recoverycases + Week4deaths + CFR, data = logdata)#0.9038

#first delete recoverycases
fit4 <- lm(Week5deaths~Activecases +Recoverycases + Week4deaths + CFR, data = logdata)#0.9114
fit4 <- lm(Week5deaths~Totalcases +Recoverycases + Week4deaths + CFR, data = logdata)#0.9078
fit4 <- lm(Week5deaths~Totalcases + Activecases + Week4deaths + CFR, data = logdata)#0.9133
fit4 <- lm(Week5deaths~Totalcases + Activecases +Recoverycases + CFR, data = logdata)#0.9113
fit4 <- lm(Week5deaths~Totalcases + Activecases +Recoverycases + Week4deaths,data = logdata)#0.9113
#second delete total
fit4 <- lm(Week5deaths~Activecases + Week4deaths + CFR, data = logdata)#0.9184
fit4 <- lm(Week5deaths~Totalcases + Week4deaths + CFR, data = logdata)#0.9112
fit4 <- lm(Week5deaths~Totalcases + Activecases + CFR, data = logdata)#0.9183
fit4 <- lm(Week5deaths~Totalcases + Activecases + Week4deaths, data = logdata)#0.9182
#Third delete week4deaths
fit4 <- lm(Week5deaths~ Week4deaths + CFR, data = logdata)#0.9132
fit4 <- lm(Week5deaths~Activecases  + CFR, data = logdata)#0.9201 use this
fit4 <- lm(Week5deaths~Activecases + Week4deaths, data = logdata)#0.9107
#fourth 
fit4 <- lm(Week5deaths~CFR, data = logdata)#0.4717
fit4 <- lm(Week5deaths~Activecases, data = logdata)#0.5951
summary(fit4)
#decide to delete recovery total week4
fit.ms2 <- lm(Week5deaths~Activecases  + CFR, data = logdata)
summary(fit.ms2)
#predict
exp(predict(fit.ms2,log(india),interval = "confidence"))
exp(predict(fit.ms2,log(india),interval = "prediction"))

```

#check collinearity and choose model
```{r}
vif(fit.ms1) #too large
vif(fit.ms2) #use this
plot(fit.ms2)
```



```{r}
gdp<-read.csv('GDP.csv',header = TRUE)
gdp<-gdp[,c("Country.Name","X2018") ]
row.names(gdp) <-  1:nrow(gdp)
levels(gdp$Country.Name) <- c(levels(gdp$Country.Name), "UK")
gdp$Country.Name[gdp$Country.Name == 'United Kingdom'] <- 'UK'
levels(gdp$Country.Name) <- c(levels(gdp$Country.Name), "USA")
gdp$Country.Name[gdp$Country.Name == 'United States'] <- 'USA'
levels(gdp$Country.Name) <- c(levels(gdp$Country.Name), "Iran")
gdp$Country.Name[gdp$Country.Name == 'Iran, Islamic Rep.'] <- 'Iran'
levels(gdp$Country.Name) <- c(levels(gdp$Country.Name), "South Korea")
gdp$Country.Name[gdp$Country.Name == 'Korea, Rep.'] <- 'South Korea'
gdp
```

```{r}
data

```


```{r}
col=data_orgin$Countries
gdp<-gdp[gdp$Country.Name %in% col,]
library(plyr)
gdp<-rename(gdp, c("Country.Name"="Countries", "X2018"="GDP"))
gdp
```


```{r}
total_origin <- merge(data_orgin,gdp,by=c("Countries"))
total_origin
india.new = total_origin[16,2:6]
india2.new = total_origin[16,c(3,4,5,8)]
india3.new= total_origin[16,c(3,4,5,6,8)]
total = total_origin[-7,]
total
```

```{r}
fit.new1 <- lm(Week5deaths~Activecases +Recoverycases + Week4deaths+GDP, data = total)
summary(fit.new1)

```


```{r}
predict(fit.new1,india2.new,interval = "confidence")
predict(fit.new1,india2.new,interval = "prediction")
```


```{r}
elder<-read.csv('Elder Percentage.csv',header = TRUE)
elder
```


```{r}

elder<-elder[,c("Country.Name","X2018") ]
row.names(elder) <-  1:nrow(elder)
levels(elder$Country.Name) <- c(levels(elder$Country.Name), "UK")
elder$Country.Name[elder$Country.Name == 'United Kingdom'] <- 'UK'
levels(elder$Country.Name) <- c(levels(elder$Country.Name), "USA")
elder$Country.Name[elder$Country.Name == 'United States'] <- 'USA'
levels(elder$Country.Name) <- c(levels(elder$Country.Name), "Iran")
elder$Country.Name[elder$Country.Name == 'Iran, Islamic Rep.'] <- 'Iran'
levels(elder$Country.Name) <- c(levels(elder$Country.Name), "South Korea")
elder$Country.Name[elder$Country.Name == 'Korea, Rep.'] <- 'South Korea'

```


```{r}
elder<-elder[elder$Country.Name %in% col,]
library(plyr)
elder<-rename(elder, c("Country.Name"="Countries", "X2018"="elder_percentage"))
elder
```


```{r}
total_origin <- merge(total_origin,elder,by=c("Countries"))
total_origin
india.new = total_origin[16,2:6]
india2.new = total_origin[16,c(3,4,5,6,8,9)]
india3.new= total_origin[16,c(3,4,5,6,8,9)]
total = total_origin[-7,]
total
india2.new
```


```{r}
fit.new2 <- lm(Week5deaths~Activecases +Recoverycases + Week4deaths+GDP+elder_percentage, data = total)
summary(fit.new2)
predict(fit.new2,india2.new,interval = "confidence")
predict(fit.new2,india2.new,interval = "prediction")
```
## final model from advance analysis
```{r}
logtot <- data.frame(total['Countries'],log(total['Totalcases']),log(total['Activecases']),log(total['Recoverycases']),log(total['Week4deaths']),log(total['CFR']),log(total['Week5deaths']),log(total['GDP']),log(total['elder_percentage']))
logtot <- logtot[-3,]
logtot
fitt <- lm(Week5deaths~Activecases  + CFR +GDP+ elder_percentage, data = logtot)
vif(fitt)
summary(fitt)
fit.final <- lm(Week5deaths~Activecases  + CFR +elder_percentage, data = logtot)
vif(fit.final)
summary(fit.final)
indiapre <- total_origin[7,c(2:11)]
indiapre
exp(predict(fit.final,log(indiapre),interval = "confidence"))
exp(predict(fit.final,log(indiapre),interval = "prediction"))
#so our final model is fit.final
```


